Laplacian envelope

### 2020-07-05 Notes/laplace.tex
pot/dep as index

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.

@@

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}}
     {\sqrt{\var(\syn_\text{ideal}\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\synid\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
    &= N\av{1}
    + N(N-1)\av{\synid^1 \syn^1(t)}^2
    - N^2\av{\synid^1 \syn^1(t)}^2 \\
    &= N(1-\av{\synid^1 \syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used $\syn^\alpha=\pm1$.
We can compute $\var(\syn_\text{ideal}\cdt\syn(\infty))$ by taking $t\to\infty$.

For the numerator, we can write
%
\begin{equation}\label{eq:overlapex}
  \av{\syn_\text{ideal}\cdt\syn(t)} = \sum_\alpha\av{\syn^\alpha_\text{ideal}\syn^\alpha(t)}
   = N \av{\syn^1_\text{ideal}\syn^1(t)},
\end{equation}
%
If the elements of $\synid$ take values $\pm1$ with probability $f\potdep$,
%
\begin{equation}\label{eq:overlapone}
  \av{\syn^1_\text{ideal}\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
\end{equation}
%
To compute this quantity, we will need to discuss how we model individual synapses.

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M\potdep$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$.
The matrices $\M\potdep$ are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
\begin{aligned}
  \M\potdep_{ij} &\geq 0, &\qquad
  \sum_j \M\potdep_{ij} &= 1.
\end{aligned}
\end{equation}
%
The upper bounds $\M\potdep_{ij} \leq 1$ follow automatically from these.

If we treat the off diagonal elements as the independent degrees of freedom, with the off diagonal elements determined by the row sum constraints (second column of \eqref{eq:constr}), the constraints are
%
\begin{equation}\label{eq:constri}
  \M\potdep_{i \neq j} \geq 0, \qquad
  \sum_{j\neq i} \M\potdep_{ij} \leq 1.
\end{equation}
%
The second of these will be referred to as the ``diagonal'' constraint.

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating/depressing is $f\potdep$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

@@

If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
If we marginalise over $n$ as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n \\
    &= \brk{ \eq \M\potdep\, \e^{rt(f\pot\M\pot+f\dep\M\dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = f\pot\M\pot + f\dep\M\dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\syn^1_\text{ideal}\syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\syn^1_\text{ideal}\syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\syn_\text{ideal}\cdt\syn(t)} - \av{\syn_\text{ideal}\cdt\syn(\infty)}
    &= N \eq \enc \, \e^{(r\frg+\hom)t} \w , \\
    \text{where} \qquad
  \enc &= f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
%We can absorb $\hom$ into $\M\potdep$ by defining
%%
%\begin{equation}\label{eq:absorbhom}
%\begin{aligned}
%  \Mh\potdep &= \M\potdep + \frac{\hom}{2rf\potdep},
%  &\qquad
%  \frgh &= f\pot\Mh\pot + f\dep\Mh\dep - \I,
%  \\&&
%  \ench &= f\pot(\Mh\pot-\I)-f\dep(\Mh\dep-\I).
%\end{aligned}
%\end{equation}
%%
%Then the SNR curve is
%%
%\begin{equation}\label{eq:SNRcurvehom}
%  \SNR(t) = \frac{\sqrt{N}  \eq \enc \,\e^{rt\frg}\w}
%                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
%\end{equation}
%%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq\w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\M\potdep$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\M\pot_{ij}}{\M\pot_{mn}} = \pdiff{\M\dep_{ij}}{\M\dep_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\M\pot_{ij}}{\M\dep_{mn}} = \pdiff{\M\dep_{ij}}{\M\pot_{mn}} = 0.
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
  \qquad
  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frg_{ij} \pm \abs{ \enc_{ij} } &\leq f\potdep, \\
  \sum_j \frg_{ij} &=0, &
  \sum_j \enc_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\M\potdep_{mn}}
     = f\potdep \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} },
\end{equation}
%
with the upper/lower sign in $\pm$ corresponding to derivatives \wrt $\M\potdep$.
The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{alignedat}{3}
    R_{ijkl} &= \
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{alignedat}
\end{equation}
%
Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
\begin{aligned}
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\pot_{kl}} &=
     (f\pot)^2   [R + R\trans + S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\pot_{ij} \partial\M\dep_{kl}} &=
     f\pot f\dep [R + R\trans - S + S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\pot_{kl}} &=
     f\pot f\dep [R + R\trans + S - S\trans]_{ijkl}, \\
  \pdiff{^2\hat{A}(s)}{\M\dep_{ij} \partial\M\dep_{kl}} &=
     (f\dep)^2   [R + R\trans - S - S\trans]_{ijkl}, \\
\end{aligned}
\end{equation}
%
This is useful in numerical maximisation.

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=0.8\linewidth]{LenvProven.eps}
  \caption[Proven envelope]{The memory curve envelope that we can prove, \eqref{eq:env}.
  The number of synapses $N=10^4$ and the number of states $M=12$.
  The dashed lines indicate the initial SNR bound, \eqref{eq:initbnd}, and the area bound, \eqref{eq:areabnd}, from \cite{Lahiri2013synapse}.}\label{fig:envproven}
\end{figure}

@@

Now we'll maximise $\snrb(\tau)$ for some fixed $\tau$, subject to the constraints \eqref{eq:constr}.
First, we will maximise the numerator of \eqref{eq:lsnr} holding $\eq\w$ fixed.
We can maximise the ration \wrt $\eq\w$ later.
Consider the Lagrangian
%
\begin{equation}\label{eq:finitelagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
        +\sum_{i j} \prn{ f\pot\mu\pot_{ij} \M\pot_{ij} + f\dep\mu\dep_{ij} \M\dep_{ij} }\\
        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
\end{aligned}
\end{equation}
%
where $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
After maximising
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:KTfinite}
  \pdiff{\CL}{\M\potdep_{ij}} = 0, \quad
  \mu\potdep_{ij} \geq 0, \quad
  \pdiff{\CL}{\mu\potdep_{ij}} \geq 0, \quad
  \mu\potdep_{ij}\pdiff{\CL}{\mu\potdep_{ij}} = 0, \quad
  \pdiff{\CL}{\lambda} = 0.
\end{equation}
%
The derivatives, for $m \neq n$ are
%
\begin{equation}\label{eq:finitederivs}
\begin{aligned}
  \frac{1}{f\potdep} \pdiff{\CL}{\Mh\potdep_{mn}} =& \,
      \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) \pm 1} \brk{\etw_m(s) - \etw_n(s)} } \\
     &+ \mu\potdep_{mn} -  \mu\potdep_{mm}
     + \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
\end{aligned}
\end{equation}
%
These have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Leibold2008serial,Ben-DayanRubin2007sparse}.

>>>>

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold~\cite{Sommer1998retrieval}.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \NNR } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \NNR &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.

@@

As discussed in \autoref{sec:recog}, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)}}
     {\sqrt{\var(\synid\cdt\syn(\infty))}},
\end{equation}
%
Where the averages and variances are over the probability distribution of the synaptic states given the sequence of plasticity events, the probability distribution of the sequence of plasticity events themselves and the probability distribution of the pattern we are testing, $\synid$.
If we also average over the time between storage and recognition, we have
%
\begin{equation}\label{eq:snrb}
  \snrb(\tau) = \intd[_0^\infty]{t} P(t\vert\tau) \snr(t)
   = \frac{1}{\tau} \intd[_0^\infty]{t} \e^{-t/\tau} \snr(t),
\end{equation}
%
This is similar to the average SNR up to time $\tau$, but with the hard cut-off replaced with a smooth exponential.
This can be expressed in terms of he Laplace transform of the SNR curve in \eqref{eq:SNRdef}:
%
\begin{equation}\label{eq:laplacedef}
  A(s) = \intd[_0^\infty]{t} \e^{-st} \snr(t),
  \qquad
  \snrb(\tau) = \frac{A(1/\tau)}{\tau}.
\end{equation}
%

First, let's look at the variances, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\synid\cdt\syn(t))
    &= \sum_{\alpha\beta} \av{\synid^\alpha \syn^\alpha(t) \synid^\beta \syn^\beta(t)}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2 \\
    &= \sum_{\alpha} \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}
    + \sum_{\alpha\neq\beta} \av{\synid^\alpha\syn^\alpha(t)}\av{\smash{\synid^\beta}\smash{\syn^\beta}(t)}
    %\\&\phantom{= \sum_{\alpha}  \av{(\synid^\alpha)^2(\syn^\alpha(t))^2}}
    - \Big(\sum_{\alpha} \av{\synid^\alpha \syn^\alpha(t)} \Big)^2  \\
    &= N\av{1}
    + N(N-1)\av{\synid^1 \syn^1(t)}^2
    - N^2\av{\synid^1 \syn^1(t)}^2 \\
    &= N(1-\av{\synid^1 \syn^1(t)}^2),
\end{aligned}
\end{equation}
%
where we used $\syn^\alpha=\pm1$.
We can compute $\var(\synid\cdt\syn(\infty))$ by taking $t\to\infty$.

For the numerator, we can write
%
\begin{equation}\label{eq:overlapex}
  \av{\synid\cdt\syn(t)} = \sum_\alpha\av{\synid^\alpha\syn^\alpha(t)}
   = N \av{\synid^1\syn^1(t)},
\end{equation}
%
If the elements of $\synid$ take values $\pm1$ with probability $f\potdep$,
%
\begin{equation}\label{eq:overlapone}
  \av{\synid^1\syn^1(t)} = f\pot \av{\syn^1(t)}_{\text{pot},t=0} - f\dep \av{\syn^1(t)}_{\text{dep},t=0}
\end{equation}
%
To compute this quantity, we will need to discuss how we model individual synapses.

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^\mu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$, 
and \(\mu\) labels the diffrerent types of plasticity event.
Potentiating events have \(\mu=\)pot, depressing events have \(\mu=\)dep.
We also use the notation
%
\begin{equation*}
  (-1)^\mu = 
    \begin{cases}
       1, & \mu = \text{pot}, \\
      -1, & \mu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices $\M^\mu$ are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \M^\mu_{ij} \geq 0 
    \quad \forall \; \mu, i, j, \qquad
  \sum_j \M^\mu_{ij} = 1
    \quad \forall \; \mu, i.
\end{equation}
%
The upper bounds $\M^\mu_{ij} \leq 1$ follow automatically from these.

If we treat the off diagonal elements as the independent degrees of freedom, with the off diagonal elements determined by the row sum constraints (second column of \eqref{eq:constr}), the constraints are
%
\begin{equation}\label{eq:constri}
  \M^\mu_{i j} \geq 0
    \quad \forall \; \mu, i, j : i \neq j, \qquad
  \sum_{j : j\neq i} \M^\mu_{ij} \leq 1
    \quad \forall \; \mu, i.
\end{equation}
%
The second of these will be referred to as the ``diagonal'' constraint.

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$ is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating/depressing is $f^\mu$, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

@@

If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
If we marginalise over $n$ as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n \\
    &= \brk{ \eq \M\potdep\, \e^{rt(f\pot\M\pot+f\dep\M\dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = \sum_\alpha f^\mu \M^\mu - \I
    = f\pot\M\pot + f\dep\M\dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\mu (-1)^\mu f^\mu (\M^\mu - \I)
    = f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq\w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\M^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\M^\mu_{ij}}{\M^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
  \qquad
  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frg_{ij} + (-1)^\mu \enc_{ij}  &\leq f^\mu, \\
  \sum_j \frg_{ij} &=0, &
  \sum_j \enc_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

Using the same approach as the area bound in \cite{Lahiri2013synapse} we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\M^\mu_{mn}}
     = f^\mu \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) + (-1)^\mu} \brk{\etw_m(s) - \etw_n(s)} }.
\end{equation}
%
The term with $\Theta(s)$ comes from differentiating $\eq$, the term with $c(s)$ comes from differentiating $\etw(s)$ and the term with $\pm1$ comes from differentiating $\enc$ in \eqref{eq:larea}.

To write down the Hessian, we will first define some notation.
We define the matrix $\mathbf{X}(s) = \fund(0) \enc \fund(s)$.
Given a four-index array $H_{ijkl}$, we define $H\trans_{ijkl} = H_{klij}$.
Then we denote
%
\begin{equation}\label{eq:hessparts}
\begin{alignedat}{3}
    R_{ijkl} &= \
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\thb_k(s) - \thb_l(s)} \\
    &&&+ \eq_i \brc{\mathbf{X}_{ik}(s) - \mathbf{X}_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_i c_i(s) \brc{\fund_{ik}(s) - \fund_{jk}(s)} 
          \brc{\etw_k(s) - \etw_l(s)},  \\
  S_{ijkl} &=
    &&\eq_i \brc{\fund_{ik}(0) - \fund_{jk}(0)} \brc{\etw_k(s) - \etw_l(s)} \\
    &&&+ \eq_k \brc{\fund_{ki}(s) - \fund_{li}(s)} \brc{\etw_i(s) - \etw_j(s)}.
\end{alignedat}
\end{equation}
%
Then we can write the Hessian as
%
\begin{equation}\label{eq:hessian}
  \pdiff{^2\hat{A}(s)}{\M^\mu_{ij} \partial\M^\nu_{kl}} =
     f^\mu f^\nu [R + R\trans + (-1)^\nu S + (-1)^\mu S\trans]_{ijkl}.
\end{equation}
%
This is useful in numerical maximisation.

@@

\begin{figure}[tb]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=0.8\linewidth]{LenvProven.eps}
  \caption[Proven envelope for the signal-to-noise ratio]
  {The memory curve envelope that we can prove, \eqref{eq:env}.
  The number of synapses $N=10^4$ and the number of states $M=12$.
  The dashed lines indicate the initial SNR bound, \eqref{eq:initbnd}, and the area bound, \eqref{eq:areabnd}, from \cite{Lahiri2013synapse}.}\label{fig:envproven}
\end{figure}

@@

Now we'll maximise $\snrb(\tau)$ for some fixed $\tau$, subject to the constraints \eqref{eq:constr}.
First, we will maximise the numerator of \eqref{eq:lsnr} holding $\eq\w$ fixed.
We can maximise the ration \wrt $\eq\w$ later.
Consider the Lagrangian
%
\begin{equation}\label{eq:finitelagrangian}
\begin{aligned}
  \CL =&\, \sum_{ij} \eq_i \enc_{ij} (\etw_i(s) - \etw_j(s))
        +\sum_{\nu i j} \prn{ f^\nu \mu^\nu_{ij} \M\pot_{ij} }
        &+ \lambda \prn{ \Delta\pr - \sum_i \eq_i\w_i },
\end{aligned}
\end{equation}
%
where $\Delta\pr$ is the constant value we are holding $\eq\w$ at.
After maximising
The Karush-Kuhn-Tucker necessary conditions for a maximum are
%
\begin{equation}\label{eq:KTfinite}
  \pdiff{\CL}{\M^\nu_{ij}} = 0, \quad
  \mu^\nu_{ij} \geq 0, \quad
  \pdiff{\CL}{\mu^\nu_{ij}} \geq 0, \quad
  \mu^\nu_{ij}\pdiff{\CL}{\mu^\nu_{ij}} = 0, \quad
  \pdiff{\CL}{\lambda} = 0.
\end{equation}
%
The derivatives, for $m \neq n$ are
%
\begin{equation}\label{eq:finitederivs}
\begin{aligned}
  \frac{1}{f^\nu} \pdiff{\CL}{\Mh^\nu_{mn}} = & \,
      \eq_m \prn{ r \brk{\thb_m(s) - \thb_n(s)}
     +\prn{rc_m(s) + (-1)^\nu} \brk{\etw_m(s) - \etw_n(s)} } \\
     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
     + \lambda \eq_m \sum_i (\fptb_{mi}(0)-\fptb_{ni}(0)) \eq_i\w_i.
\end{aligned}
\end{equation}
%
These have the same form as for $s=0$, which allowed us to argue that the model with maximal area must have the serial topology, \cite{Leibold2008serial,Ben-DayanRubin2007sparse}.

The one missing ingredient is the scale invariance.
Define the scale transformation:
%
\begin{equation}\label{eq:scale}
  \lambda * \M = (1-\lambda)\I + \lambda \M.
\end{equation}
%
The Laplace transform has the following scaling property
%
\begin{equation}\label{eqlaplacescale}
  A(\lambda s;\lambda * \M^\nu) = A(s;\M^\nu).
\end{equation}
%
At $s=0$, it is scale invariant.
This means that we can take a pair of matrices that violates the ``diagonal'' constraints (second inequality of \eqref{eq:constri}) and use the scale transform \eqref{eq:scale} to construct a pair of matrices that satisfy the ``diagonal'' constraints and have the same area.
This allowed us to ignore the diagonal constraints.
We can't do that for $s>0$.
